---
title: "R Notebook"
author: eyan
date: 23/01/2021
output: rmarkdown::github_document
---
```{r}
library('tidyverse')
library('data.table')
```
```{r}
url <- 'http://bit.ly/CarreFourDataset'
sup_market <- read.csv(url)
head(sup_market)
```
```{r}
tail(sup_market)
```
```{r}
colSums(is.na(sup_market))
```
No null values were found
```{r}
dim(sup_market)
sup_market[duplicated(sup_market)]
```
No duplicates
```{r}
summary(sup_market)
```

# Reducing the number of features using PCA
```{r pca}
# importing the necessary library for PCA
library('FactoMineR')
library('ggbiplot')
```
```{r}
label_encode <- function (column,dataframe){
  data <- dataframe[[column]]
  new_data <- as.integer(factor(data))
  # print(new_data)
  return (new_data)
}
supermarket.org <- sup_market[,-c(1,9,10)]

```
```{r}
# Label encoding columns
names_ <- c('Branch','Customer.type','Gender','Product.line','Payment')
for(nm in names_){
  supermarket.org[nm] <- label_encode(nm,supermarket.org)
}
supermarket <- supermarket.org
head(supermarket)
```
```{r}
# Geting the principle components from the supermarkets dataset.
results <- prcomp(supermarket[,-c(10)], center = TRUE, scale. = TRUE)
summary(results)
```
The first principal component has the largest proportion explaining 41% of the variance in the data.
The first 9 components explain the all variance in the data, making the last 3 obsolete
```{r}
screeplot(results, type = "l", npcs = 13, main = "Screeplot of the first 12 Principal components")
abline(h = 0.9, col="red", lty=5)
```
The first 9 principal components explain the most in the data, explainig 99% of the variance. These are the most important features in the dataset.

```{r}
ggbiplot(results,labels.size = 5, alpha = 0.5)
```
Unit price,gross income,tax and quantity contribute to PC1
Payment,gender,Product line and customer type contribute to PC2
```{r}
clusters <- function (name,pc1=1,pc2=2){
  cl_list <- factor(sup_market[[name]])
  # species <- cl_list$species
  ggbiplot(results,,choices=c(pc1,pc2),ellipse = TRUE,groups = cl_list, obs.scale = 1, var.scale = 1,labels.size = 5, alpha = 0.5)
}
clusters('Gender')
```
The clusters of the gender show:
  Males are clustered at the top characterised by all features but the branch, while females are clustered at the bottom with unitprice,tax,gross income,rating and branch influencing their buying.
```{r}
clusters('Gender',3,4)
```
Since the 3rd and 4th princpal components explain lower variances than the first two, i doesnt do that good a job at showing the pattern shown in the previous plot. although it does give a hint on where the different genders lie.
```{r}
clusters('Gender',11,12)
```
The last 2 principle components perform even worse since they ecplain so little of the variance in the dataset.


# Reducing the attributes using Feature Selection
## Method 1 filter method: high correlation filter
```{r}
library('caret')
library('corrplot')
```
```{r}
supermarket_ <- supermarket.org[,-c(10)]
corr_ <- cor(supermarket_)
corr_
```
High correlations from tax,gross income ,cogs and total
```{r}
high_cor <- findCorrelation(corr_, cutoff=0.75)
high_cor
```

```{r}
# Removing the highly correlated features in index (9,12,7)
final <- supermarket_[-high_cor]
```
```{r}
# Plotting the correlations for comparison
par(mfrow = c(1,2))
corrplot(corr_, order = "hclust")
corrplot(cor(final), order = "hclust")
```
Feature selection removes the most correlated features from the dataset in this case were left with 9 features that are not overly correlated

## Method 2 : Wrapper Methods
```{r}
library('clustvarsel')
library('mclust')
```
```{r}
# Sequential forward greedy search using the default number of groups
out <- clustvarsel(supermarket_)
out
```
```{r}
final2 <- supermarket_[,out$subset]
head(final2)
```
Using clustvarsel 3 groups were taken as important in creating the clusters of the dataset
```{r}
mod <- Mclust(final2)
summary(mod)
```
There were achieved 7 clusters in the selected subset of data
```{r}
plot(mod,c("classification"))
```
The unsupervised clustering shows 7 clusters from the 3 attributes selected